Imtiyaaz

### 2022-11-16 src/All variables.Rmd
Update All variables.Rmd

```{r}
eval_results <- function(true, predict) {
  actual <- data.matrix(true)
  SSE <- sum((predict - actual)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE / SST
  data.frame(
    MSE = MSE(predict, true),
    MAE = MAE(predict, true),
    RMSE = RMSE(predict, true),
    MAPE = MAPE(predict, true),
    Rsquare = R_square
  )
}
```

```{r}
df <-read.csv('../data/countries1.csv')
df
subset1 <- df %>% select("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
subset1
```

```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

```{r}
lm.slr <- lm(pov ~ income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
summary(lm.slr)

```
```{r}
subset1$income <- factor(subset1$income)
subset1$income <- as.numeric(subset1$income)

@@

```{r}
str(subset1)
df0 <- subset1
scaler <- apply(X = df0, MARGIN = 2, FUN = sd)
subset1 <- as.data.frame(apply(df0, 2, function(x) x/sd(x, na.rm = TRUE)))
subset1
apply(subset1, 2, sd)

```
```{r}
model1 <- lm(pov ~ ., data = subset1)
summary(model1)
vif(model1)

```
```{r}
subset2 <- subset(subset1,select =c("pov","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
subset2 <- na.omit(subset2)
train.x <- data.matrix(subset(subset2, select = -pov))
train.y = data.matrix(subset2$pov)
```

```{r}
corrplot(cor(subset2, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
set.seed(123)
rate <- 0.8
train.size <- round(nrow(subset2)*rate)
sample(sample(nrow(subset2), train.size))
training <- subset2[sample, ]
test <- subset2[-sample, ]

```
```{r}
model3 <- lm(pov ~., data = training)
summary(model3)

@@

```{r}
set.seed(123)
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x,
                    train.y,
                    alpha = 0,
                    lambda = cv_ridge$lambda.min)
t(coef(glm_ridge))

```{r}
train.x <- model.matrix(pov ~., data = training)[, -1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```

```{r}
model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
t(coef(model_ridge_trial1))
```

```{r}
model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
t(coef(model_ridge_trial2))
```

```{r}
model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
t(coef(model_ridge_trial3))
```

```{r}
table_summary <- cbind(coef(model_ridge_trial1),
                       coef(model_ridge_trial2),
                       coef(model_ridge_trial3))

colnames(table_summary) <- c("lambda = 0.1",
                             "lambda = 0.5",
                             "lambda = 1")

table_summary
```

```{r}
set.seed(123)
cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
glm_lasso <- glmnet(train.x,train.y, 
                    alpha = 1, lambda = cv_lasso$lambda.min)
t(coef(glm_lasso))
```

```{r}
# Obtain performance of MLR model
fit <- model1$fitted.values
true <- subset2$pov
summary_model1 <- eval_results(true, fit)

# Obtain performance of ridge model
fit <- predict(glm_ridge, newx = train.x)
true <- train.y
summary_ridge <- eval_results(true, fit)

# Obtain performance of LASSO model
fit <- predict(glm_lasso, newx = train.x)
true <- train.y
summary_lasso <- eval_results(true, fit)

# Combine all performances into a single output
summary <- rbind(summary_model1, summary_ridge, summary_lasso)
rownames(summary) <- c("MLR", "Ridge_model", "LASSO_model" )
colnames(summary)[5] <- "R^2"
knitr::kable(summary, digits = 3)

```{r}
knitr::kable(summary_MLR, digits = 3)
knitr::kable(summary_ridge[,-5], digits = 3)
```

knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)
```

>>>>

First we will write the function "eval_results(true,predict) to evaluate the models
```{r}
eval_results <- function(true, predict) {
  actual <- data.matrix(true)
  SSE <- sum((predict - actual)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE / SST
  data.frame(
    MSE = MSE(predict, true),
    MAE = MAE(predict, true),
    RMSE = RMSE(predict, true),
    Rsquare = R_square
  )
}
```
First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
```{r}
df <-read.csv('../data/countries1.csv')
df
subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
subset1
```

Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
```{r}
lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
summary(lm.slr)

```
Next we will be perfoming regularisation. We have to use this function as not all the varibales are numerical. Therefore to overcome it we build the function below.
```{r}
subset1$income <- factor(subset1$income)
subset1$country.code <- factor(subset1$country.code)

sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

  return(sd(vct, na.rm = T))
}

std0 <- function(vct, scl) {
  if (!is.numeric(vct)) {
    return(vct)
  }

  return(vct/scl)
}

scale <- function(data) {
  unlist(lapply(data, sd0))
}

standardise <- function(data) {
  scaler <- scale(data)

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}
```

Next we will perform standardistation to the dataset.
```{r}
str(subset1)
df0 <- subset1
subset1
apply(subset1, 2, sd)

@@

Next we will plot a pov vs income scatter plot to see the effect of income on pov. From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov. This is consistent with our hypothesis that countries that are richer do not have as much poverty issue as the poorer countries.
```{r}
ggplot(data = subset1, aes(x = income, y = pov)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Pov vs. Income ",
       x = "income",
       y = "pov")

```{r}
model1 <- lm(pov ~ ., data = subset1)
summary(model1)
vif(model1)

```
```{r}
subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
subset2 <- na.omit(subset2)
subset2
train.x <- data.matrix(subset(subset2, select = -pov))
train.y = data.matrix(subset2$pov)
```

```{r}
split_train_test <- function(data) {
set.seed(1984)
# split data
seeds <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() == 1)

plants <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() != 1)

isComplete <- which(complete.cases(plants))
idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))

training <- plants[-idx,] %>% 
  rbind(seeds)
test <- plants[idx,]

return(list(training, test))
}
```

```{r}

training <- split_train_test(subset2)[[1]]
test <- split_train_test(subset2)[[2]]
```

```{r}
ntrain <- 1:nrow(training)
combine <- rbind(training, test)
combine.mtrx <- model.matrix(pov ~ ., data = combine)
train.mtrx <- combine.mtrx[ntrain,]
test.mtrx <- combine.mtrx[-ntrain,]

train.x <- train.mtrx[,-1]
train.y <- training$pov

test.x <- test.mtrx[,-1]
text.y <- test$pov
```

```{r}
model3 <- lm(pov ~., data = training)
summary(model3)

@@
